Supplementary information

### 2018-06-13 manuscript/02_methods.Rmd
Update reviewer response
Include sample size analysis.

Fix other tables.

Do some text revisions.

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,max}$, we only used values already reported in TRY as being at 25°C.
For $R_{d}$, we normalized the values to 25°C based on reported leaf temperature values following the same methods as @atkin_global_2015.
For $J_{max}$, we normalized the values to 25°C based on reported leaf temperature values using the temperature response function described in @kattge_2007_temperature (Equation 1 therein).
To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

@@

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows:
For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
the missing values $\mathbf{x'}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

@@

>>>>

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,max}$, we only used values already reported in TRY as being at 25°C.
For $R_{d}$, we normalized the values to 25°C based on reported leaf temperature values following the same methods as @atkin_global_2015.
For $J_{max}$, we normalized the values to 25°C based on reported leaf temperature values using the temperature response function described in @kattge_2007_temperature (Equation 1 therein).
To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled, where at least one of the above traits was observed, and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

@@

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
Therefore, our algorithm also included multiple imputation of partially missing data, which proceeded as follows:
For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
the missing values $\mathbf{x'}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

@@

We emphasize that the imputation of missing data is performed iteratively as parameters are being estimated, such that each set of imputed values is conditioned on the current sampled state of the parameters.
This approach is therefore distinct from single imputation, where data are imputed first in a separate step prior to parameter estimation.
A demonstration of this multiple imputation approach and how it is used to estimate trait covariance is provided in a literate programming vignette in the supplementary information to this manuscript.

### 2018-06-13 supplementary_information.Rmd
Update reviewer response
Include sample size analysis.

Fix other tables.

Do some text revisions.

```{r libraries, include = FALSE}
library(knitr)
library(kableExtra)
library(drake)
```

```{r sample_size_table, echo = FALSE, results = "asis"}
kable(
  readd(sample_size_wide),
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 1, 5) %>%
  group_rows("Needleleaf", 6, 8) %>%
  group_rows("Shrub", 9, 11) %>%
  group_rows("Grass", 12, 14)
```

```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
kable(
  readd(corr_processed), "latex",
  booktabs = TRUE, longtable = TRUE, escape = FALSE,
  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)"),
  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
)
```

>>>>

```{r libraries, include = FALSE}
library(knitr)
library(kableExtra)
library(drake)
library(ggplot2)
```

```{r sample_size_table, echo = FALSE, results = "asis"}
ssw <- readd(sample_size_wide)
kable(
  ssw,
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 2, 6) %>%
  group_rows("Needleleaf", 7, 9) %>%
  group_rows("Shrub", 10, 12) %>%
  group_rows("Grass", 13, 15) %>%
  landscape()
```

```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
corr_results <- readd(corr_processed)
kable(
  corr_results %>%
    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
  "latex",
  booktabs = TRUE, longtable = TRUE, escape = FALSE,
  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)", "present", "missing"),
  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
)

correlation_scatter_cap <- paste(
  "Pairwise mean squared trait correlation values as a function of sample size.",
  "Note that the x-axis uses a $\\log_10$ scale."
)
correlation_fig_cap <- paste(
  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
  "as a function of sample size for each plant functional type."
)
```

```{r correlation_sample_size_pairs, echo = FALSE, fig.cap = correlation_scatter_cap}
corr_results %>%
  mutate(
    significant = factor(significant, c("TRUE", "FALSE")) %>%
      forcats::lvls_revalue(c("yes", "no"))
  ) %>%
  ggplot() +
  aes(x = present + 0.1, y = Mean, color = significant) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ pft) +
  scale_x_log10() +
  labs(x = "Sample size", y = "Correlation") +
  theme_bw() +
  theme(
	legend.position = "bottom"
  )
```

```{r correlation_sample_size_analysis, echo = FALSE, fig.cap = correlation_fig_cap}
corr_binned <- corr_results %>%
  dplyr::mutate(
    ss_bin = cut(
      present,
      c(0, 1, 10, 25, 50, 100, 500, 1000, Inf),
      c("0", "< 10", "< 25", "< 50", "< 100", "< 500", "< 1000", "> 1000"),
      include.lowest = TRUE,
      right = FALSE
    )
  ) %>%
  group_by(pft, ss_bin) %>%
  summarize(frac_significant = mean(significant))

ggplot(corr_binned) +
  aes(x = ss_bin, y = frac_significant) +
  geom_segment(aes(xend = ss_bin, yend = 0)) +
  geom_point() +
  facet_wrap(~ pft) +
  labs(x = "Sample size", y = "Fraction of significant correlations") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.9))
```
